A Systematic Literature Review on Extensions to Role Based Access Control

### 2013-04-25 paper/sections/categorization.tex
Adding some edits to the discussion section.

  \item Context: The extension model integrates contextual information into the RBAC standard model. Context is defined as a user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.

  \item Constraint (Const): The extension model provides conditional restrictions on permissions of given roles. The constraint is either static or dynamic. For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician. This example describes a doctor's permission with conditional restriction - ``only when the doctor is assigned as the designated primary care physician can the doctor modify a particular medical record''.

  \item Organizational (Org): The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations. Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.

  \item Privacy (Priv): The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.

  \item Task: The extension model provides a task entity which are associated with permissions and roles. A task is a fundamental unit of a business activity. Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions. Roles are first associated with tasks, which are associated permissions. For example, the employee role is associated with a task, which is to write a report. Then, this task is associated with a permission.

  \item Spatio-Temporal: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies. For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.

  \item Spatial: The extension model provides spatial constraints, location-based constraints in specifying access
	control policies. For example, in organizations, locations are enforced while a
	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
	In such cases, the part-time employee role should access required resources only when the user is in the location. 
	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.

  \item Temporal (Temp):  The extension model provides temporal constraints, time-based constraints in specifying access
	control policies. For example, in organizations, periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.

>>>>

  \item \textbf{Context}: The extension model integrates contextual information into the RBAC standard model. Context is defined as a user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.

  \item \textbf{Constraint (Const)}: The extension model provides conditional restrictions on permissions of given roles. The constraint is either static or dynamic. For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician. This example describes a doctor's permission with conditional restriction - ``only when the doctor is assigned as the designated primary care physician can the doctor modify a particular medical record''.

  \item \textbf{Organizational (Org)}: The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations. Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.

  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.

  \item \textbf{Task}: The extension model provides a task entity which are associated with permissions and roles. A task is a fundamental unit of a business activity. Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions. Roles are first associated with tasks, which are associated permissions. For example, the employee role is associated with a task, which is to write a report. Then, this task is associated with a permission.

  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies. For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.

  \item \textbf{Spatial}: The extension model provides spatial constraints, location-based constraints in specifying access
	control policies. For example, in organizations, locations are enforced while a
	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
	In such cases, the part-time employee role should access required resources only when the user is in the location. 
	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.

  \item \textbf{Temporal (Temp)}:  The extension model provides temporal constraints, time-based constraints in specifying access
	control policies. For example, in organizations, periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.

### 2013-04-25 paper/sections/discussion.tex
Adding some edits to the discussion section.

\section{Discussion} \label{sec:discussion}
%\subsection{Find appropriate extended Role-based access control models for a system}

This section discusses two issues: first one is about evaluation of extended role-based access control models and second one is to implement security system based on access control models.

\subsection{Metrics for Extended Access Control Models Evaluation }

Researchers provide metrics for access control system evaluation or security system in general~\cite{hu2012:NIST7874}. They collected metrics, which can be applicable for not only access control models, but also requirements and system implementation, and extended application such as verification and testing. From the metrics, we select metrics that can be applicable for evaluation of extended RBAC models as follows:

\begin{itemize}
	\item \textbf{Ease of Privilege Assignments}: this metric measures number of steps for assigning, changing, deleting a privilege for users and roles when an end-user or administrators manage RBAC. If one takes more steps for such a task, it may give more chance to make human or system errors due more steps. If you takes less step, it may reduce to make mistakes.  While extended RBAC may has similar features since from RBAC, we measure worst cases after extended, and gap between these two models.

	\item \textbf{Syntactic and semantic support for specifying AC rules}: this metric checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints. This support helps specify complex and flexible access controls. For example, given a spatial RBAC model, one may measure whether spatial constraints can be specified using various logic expressions.

	\item \textbf{Delegation of administrative capabilities}: this metric measure how easy and secure when an access control model supports for privilege delegation of administrators to another user.

	\item \textbf{Capable of combining different policy models}: this metric measures whether an access model has a feature to combine more than two policies.

	\item \textbf{Capable of conflict resolution}: These metric checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.

	\item \textbf{Least privilege principle support}: This metric checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning. Least privilege principle support can reduce damage when malicious users gain legitimate privileges.

	\item \textbf{Separation of Duty (SoD)}: This metric checks whether an access model supports for static SoD and/or dynamic SoD. Note that NIST Standard supports both static SoD and dynamic SoD described in Section~\ref{sec:core-rbac}.
\end{itemize}

One can measure these metrics for a given extended access control model. These metrics focus on comparing specific features support of access control models.

\subsection{Extended RBAC Adoption}

>>>>

\section{Discussion} \label{sec:discussion}

The research questions we identified, presented results for and analyzed provide specific views of portions of the extension model landscape. 
By stepping back, and looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
We break the discussion into two issues: recommendations for the enhancement of the evaluation of extensions to role-based access control models and guidelines for developers needing to choose an extension model to meet their requirements.

\subsection{Metrics for the Evaluation of Extensions to RBAC}

A small percentage of primary sources presented substantial evaluations or implementations of their model in practice.  This in turn leaves researchers and practioners little data from which to compare one model to the next. And given the wide swath of popluar and critical domains being targeted by RBAC extension models, we propose a set of metrics by which RBAC extension models should be evaluated. Hu, Vincent C. and Scarfone, Karen provide metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}. Hu, Vincent C. and Scarfone, Karen collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing. From these metrics, and guided by the data surrounding how extension models are currently evaluated, we select metrics that can be applicable for evaluation of extended RBAC models as follows:

\begin{itemize}
	\item \textbf{Ease of Privilege Assignments}: measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manage RBAC. If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity. While extended RBAC may have similar features to that of the RBAC standard model, we measure worst cases after extended, and the gap between an extension model and the RBAC standard model.

	\item \textbf{Syntactic and semantic support for specifying AC rules}: checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints. This support helps specify complex and flexible access controls. For example, given a spatial RBAC model, one may measure whether spatial constraints can be specified using various logic expressions.

	\item \textbf{Delegation of administrative capabilities}: measures how easy and secure an access control model supports for privilege delegation of administrators to another user.

	\item \textbf{Capable of combining different policy models}: measures whether an access model has a feature to combine more than two policies.

	\item \textbf{Capable of conflict resolution}: checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.

	\item \textbf{Least privilege principle support}: checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning. Least privilege principle support can reduce damage when malicious users gain legitimate privileges.

	\item \textbf{Separation of Duty (SoD)}: checks whether an access model supports for static SoD and/or dynamic SoD. Note that NIST Standard supports both static SoD and dynamic SoD described in Section~\ref{sec:core-rbac}.
\end{itemize}

We recommend these metrics be incorporated in any extension models presented hence forth. A future work could include providing metric measurements for the primary sources prsented here. 
A developer could attemot to apply these metrics to an extension model not already evaluated when attempting to choose a model. 
These metrics focus on comparing specific features support of access control models.

\subsection{Adoption of Extensions to RBAC}
